---
title: "Stat 566 - Project"
subtitle: "H1N1 Vaccination"
author: "Natalie Chisam, Rachel Flodin, Kaelan Yu"
date: "4/13/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(formatR)
knitr::opts_chunk$set(echo=FALSE,tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

```{r data}
# Source: https://www.kaggle.com/arashnic/flu-data
# Import Data
H1N1.dat <- read.csv("./H1N1_Flu_Vaccines.csv")

# Look at the dimensions
dim(H1N1.dat)
```

```{r introData}
# Look at a summary
summary(H1N1.dat)

# Look at the data structure
str(H1N1.dat)
```

```{r variableConv}
# Add "NA" to character variables
H1N1.dat$age_group[H1N1.dat$age_group==""] <- NA
H1N1.dat$education[H1N1.dat$education==""] <- NA
H1N1.dat$race[H1N1.dat$race==""] <- NA
H1N1.dat$sex[H1N1.dat$sex==""] <- NA
H1N1.dat$income_poverty[H1N1.dat$income_poverty==""] <- NA
H1N1.dat$marital_status[H1N1.dat$marital_status==""] <- NA
H1N1.dat$rent_or_own[H1N1.dat$rent_or_own==""] <- NA
H1N1.dat$employment_status[H1N1.dat$employment_status==""] <- NA
H1N1.dat$census_msa[H1N1.dat$census_msa==""] <- NA

# Binary variables - convert to factors with Yes/No labels
H1N1.dat$h1n1_vaccine <- factor(H1N1.dat$h1n1_vaccine, levels=c(0,1), labels=c("No", "Yes"), exclude=NULL)
H1N1.dat$h1n1_vaccine <- addNA(H1N1.dat$h1n1_vaccine)
H1N1.dat$seasonal_vaccine <- factor(H1N1.dat$seasonal_vaccine, levels=c(0,1), labels=c("No", "Yes"), exclude=NULL)
H1N1.dat$seasonal_vaccine <- addNA(H1N1.dat$seasonal_vaccine)
H1N1.dat$behavioral_wash_hands <- factor(H1N1.dat$behavioral_wash_hands, levels=c(0,1), labels=c("No", "Yes"), exclude=NULL)
H1N1.dat$behavioral_wash_hands <- addNA(H1N1.dat$behavioral_wash_hands)
H1N1.dat$behavioral_large_gatherings <- factor(H1N1.dat$behavioral_large_gatherings, levels=c(0,1), labels=c("No", "Yes"), exclude=NULL)
H1N1.dat$behavioral_large_gatherings <- addNA(H1N1.dat$behavioral_large_gatherings)
H1N1.dat$behavioral_antiviral_meds <- factor(H1N1.dat$behavioral_antiviral_meds, levels=c(0,1), labels=c("No", "Yes"), exclude=NULL)
H1N1.dat$behavioral_antiviral_meds <- addNA(H1N1.dat$behavioral_antiviral_meds)
H1N1.dat$behavioral_avoidance <- factor(H1N1.dat$behavioral_avoidance, levels=c(0,1), labels=c("No", "Yes"), exclude=NULL)
H1N1.dat$behavioral_avoidance <- addNA(H1N1.dat$behavioral_avoidance)
H1N1.dat$behavioral_face_mask <- factor(H1N1.dat$behavioral_face_mask, levels=c(0,1), labels=c("No", "Yes"), exclude=NULL)
H1N1.dat$behavioral_face_mask <- addNA(H1N1.dat$behavioral_face_mask)
H1N1.dat$behavioral_outside_home <- factor(H1N1.dat$behavioral_outside_home, levels=c(0,1), labels=c("No", "Yes"), exclude=NULL)
H1N1.dat$behavioral_outside_home <- addNA(H1N1.dat$behavioral_outside_home)
H1N1.dat$behavioral_touch_face <- factor(H1N1.dat$behavioral_touch_face, levels=c(0,1), labels=c("No", "Yes"), exclude=NULL)
H1N1.dat$behavioral_touch_face <- addNA(H1N1.dat$behavioral_touch_face)
H1N1.dat$doctor_recc_h1n1 <- factor(H1N1.dat$doctor_recc_h1n1, levels=c(0,1), labels=c("No", "Yes"), exclude=NULL)
H1N1.dat$doctor_recc_h1n1 <- addNA(H1N1.dat$doctor_recc_h1n1)
H1N1.dat$doctor_recc_seasonal <- factor(H1N1.dat$doctor_recc_seasonal, levels=c(0,1), labels=c("No", "Yes"), exclude=NULL)
H1N1.dat$doctor_recc_seasonal <- addNA(H1N1.dat$doctor_recc_seasonal)
H1N1.dat$chronic_med_condition <- factor(H1N1.dat$chronic_med_condition, levels=c(0,1), labels=c("No", "Yes"), exclude=NULL)
H1N1.dat$chronic_med_condition <- addNA(H1N1.dat$chronic_med_condition)
H1N1.dat$child_under_6_months <- factor(H1N1.dat$child_under_6_months, levels=c(0,1), labels=c("No", "Yes"), exclude=NULL)
H1N1.dat$child_under_6_months <- addNA(H1N1.dat$child_under_6_months)
H1N1.dat$health_worker <- factor(H1N1.dat$health_worker, levels=c(0,1), labels=c("No", "Yes"), exclude=NULL)
H1N1.dat$health_worker <- addNA(H1N1.dat$health_worker)
H1N1.dat$health_insurance <- factor(H1N1.dat$health_insurance, levels=c(0,1), labels=c("No", "Yes"), exclude=NULL)
H1N1.dat$health_insurance <- addNA(H1N1.dat$health_insurance)

# Convert other variables to factors
H1N1.dat$age_group <- addNA(H1N1.dat$age_group)
H1N1.dat$education <- addNA(H1N1.dat$education)
H1N1.dat$race <- addNA(H1N1.dat$race)
H1N1.dat$sex <- addNA(H1N1.dat$sex)
H1N1.dat$income_poverty <- addNA(H1N1.dat$income_poverty)
H1N1.dat$marital_status <- addNA(H1N1.dat$marital_status)
H1N1.dat$rent_or_own <- addNA(H1N1.dat$rent_or_own)
H1N1.dat$employment_status <- addNA(H1N1.dat$employment_status)
H1N1.dat$h1n1_concern <- addNA(H1N1.dat$h1n1_concern)
H1N1.dat$h1n1_knowledge <- addNA(H1N1.dat$h1n1_knowledge)
H1N1.dat$opinion_h1n1_vacc_effective <- addNA(H1N1.dat$opinion_h1n1_vacc_effective)
H1N1.dat$opinion_h1n1_risk <- addNA(H1N1.dat$opinion_h1n1_risk)
H1N1.dat$opinion_h1n1_sick_from_vacc <- addNA(H1N1.dat$opinion_h1n1_sick_from_vacc)
H1N1.dat$opinion_seas_vacc_effective <- addNA(H1N1.dat$opinion_seas_vacc_effective)
H1N1.dat$opinion_seas_risk <- addNA(H1N1.dat$opinion_seas_risk)
H1N1.dat$opinion_seas_sick_from_vacc <- addNA(H1N1.dat$opinion_seas_sick_from_vacc)
H1N1.dat$census_msa <- addNA(H1N1.dat$census_msa)
H1N1.dat$household_adults <- addNA(H1N1.dat$household_adults)
H1N1.dat$household_children <- addNA(H1N1.dat$household_children)
```

## Variables Removed

**respondent_id** – unique identifier, ranges from 0 to 26706

**hhs_geo_region** - Respondent's residence using a 10-region geographic classification defined by the U.S. Dept. of Health and Human Services. Values are represented as short random character strings.

**employment_industry** - Type of industry respondent is employed in. Values are represented as short random character strings.

**employment_occupation** - Type of occupation of respondent. Values are represented as short random character strings.


```{r removeVars}
# If the dataset still has all 38 columns, remove the following variables:
# respondent_id,hhs_geo_region,employment_industry,employment_occupation 
if(ncol(H1N1.dat)==38)
{
H1N1.dat <- H1N1.dat[,-c(1,31,35,36)]
}
```

## Descriptions of the variables

**h1n1_vaccine** - Whether respondent received H1N1 flu vaccine.

**seasonal_vaccine** - Whether respondent received seasonal flu vaccine.

**age_group** - Age group of respondent.

**education** - Self-reported education level.

**race** - Race of respondent.

**sex** - Sex of respondent.

**income_poverty** - Household annual income of respondent with respect to 2008 Census poverty thresholds.

**marital_status** - Marital status of respondent.

**rent_or_own** - Housing situation of respondent.

**employment_status** - Employment status of respondent.

**h1n1_concern** - Level of concern about the H1N1 flu.
0 = Not at all concerned; 1 = Not very concerned; 2 = Somewhat concerned; 3 = Very concerned.

**h1n1_knowledge** - Level of knowledge about H1N1 flu.
0 = No knowledge; 1 = A little knowledge; 2 = A lot of knowledge.

**behavioral_wash_hands** - Has frequently washed hands or used hand sanitizer. (binary)

**behavioral_large_gatherings** - Has reduced time at large gatherings. (binary)

**behavioral_antiviral_meds** - Has taken antiviral medications. (binary)

**behavioral_avoidance** - Has avoided close contact with others with flu-like symptoms. (binary)

**behavioral_face_mask** - Has bought a face mask. (binary)

**behavioral_outside_home** - Has reduced contact with people outside of own household. (binary)

**behavioral_touch_face** - Has avoided touching eyes, nose, or mouth. (binary)

**doctor_recc_h1n1** - H1N1 flu vaccine was recommended by doctor. (binary)

**doctor_recc_seasonal** - Seasonal flu vaccine was recommended by doctor. (binary)

**chronic_med_condition** - Has any of the following chronic medical conditions: asthma or an other lung condition, diabetes, a heart condition, a kidney condition, sickle cell anemia or other anemia, a neurological or neuromuscular condition, a liver condition, or a weakened immune system caused by a chronic illness or by medicines taken for a chronic illness. (binary)

**child_under_6_months** - Has regular close contact with a child under the age of six months. (binary)

**health_worker** - Is a healthcare worker. (binary)

**health_insurance** - Has health insurance. (binary)

**opinion_h1n1_vacc_effective** - Respondent's opinion about H1N1 vaccine effectiveness.
1 = Not at all effective; 2 = Not very effective; 3 = Don't know; 4 = Somewhat effective; 5 = Very effective.

**opinion_h1n1_risk** - Respondent's opinion about risk of getting sick with H1N1 flu without vaccine.
1 = Very Low; 2 = Somewhat low; 3 = Don't know; 4 = Somewhat high; 5 = Very high.

**opinion_h1n1_sick_from_vacc** - Respondent's worry of getting sick from taking H1N1 vaccine.
1 = Not at all worried; 2 = Not very worried; 3 = Don't know; 4 = Somewhat worried; 5 = Very worried.

**opinion_seas_vacc_effective** - Respondent's opinion about seasonal flu vaccine effectiveness.
1 = Not at all effective; 2 = Not very effective; 3 = Don't know; 4 = Somewhat effective; 5 = Very effective.

**opinion_seas_risk** - Respondent's opinion about risk of getting sick with seasonal flu without vaccine.
1 = Very Low; 2 = Somewhat low; 3 = Don't know; 4 = Somewhat high; 5 = Very high.

**opinion_seas_sick_from_vacc** - Respondent's worry of getting sick from taking seasonal flu vaccine.
1 = Not at all worried; 2 = Not very worried; 3 = Don't know; 4 = Somewhat worried; 5 = Very worried.

**census_msa** - Respondent's residence within metropolitan statistical areas (MSA) as defined by the U.S. Census.

**household_adults** - Number of other adults in household, top-coded to 3.

**household_children** - Number of children in household, top-coded to 3.


```{r EDA, fig.height=5, fig.width=8}
for(i in colnames(H1N1.dat)){
  plot(H1N1.dat[,i], main=i, names.arg=levels(H1N1.dat[,i]), cex.names=.7)
}
```
